---
title: "ST558_HW7"
format: html
editor: visual
author: Evan Whitfield, Ralph Church, Ryan Friedman
---

### Reading in the Data

First, load the libraries needed for importing and working with the dataset.

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidymodels)
```

Next, read the CSV file into R using the `read_csv()` function from the **readr** package.

```{r}
bike <- read_csv(
  "Data/SeoulBikeData.csv",
  locale = locale(encoding = "Windows-1252"),
  show_col_types = FALSE
) |>
  clean_names()

# convert few of variables to factors
bike <- bike |>
  mutate(
    date = as.Date(date, format = "%d/%m/%Y"),
    across(
      c(seasons, holiday, functioning_day),
      as.factor
    )
  )
```

Finally, inspect the data to verify that it was imported correctly.

```{r}
head(bike)
glimpse(bike)
summary(bike)
```

The dataset is stored in the object `bike`, which will be used throughout the analysis. The `head()` function displays the first few rows, `glimpse()` provides the structure and variable types, and `summary()` gives descriptive statistics for each variable.

### Check for missing values

```{r}
missing_values <- data.frame(
  Variable = names(bike),
  Missing_Values = colSums(is.na(bike))
)

missing_values
```

The dataset has no missing values, so no data needed to be removed or filled in before the analysis.

### Split the training set

```{r}
set.seed(10)

bike_split <- initial_split(bike, prop = 0.70)

bike_train <- training(bike_split)
bike_test <- testing(bike_split)

bike_train
bike_test
```

Next, verify the size of each dataset using **dim()**

```{r}
dim(bike_train)
dim(bike_test)
```
