---
title: "HW7_RF"
format: html
---

```{r}
library(tidyverse)
library(tidymodels)
library(lubridate)
library(janitor)
library(gt)
```

## Read in the data

```{r}
# adding locale= to fix Error in nchar(x, "width") : invalid multibyte string, element 1
bike <- read_csv("../Data/SeoulBikeData.csv", locale=locale(encoding="latin1"))
bike
```

## Conduct Basic EDA

- Begin with `str()` and `summary()` to examine variables

```{r}
str(bike)
```

```{r}
summary(bike)
```
- Notes on the above:

  - 14 variables.  Some numeric and some categorical. Numeric are correctly coming in as numbers
  
  - 8760 rows = 365 days x 24 hours
  
  - Categorical variable unique values: 365 dates, 4 seasons, 2 (Y/N) for holiday, 2 (Y/N) for functioning day

- Fix variable names

```{r}
colnames(bike) <- c("date", "bike_count", "hour", "temp", "hum", "wind", "vis", "dewpt", "solrad", "rain", "snow", "season", "holiday", "functioning")
colnames(bike)
```

- Fix the date and turn the appropriate variables into factors

```{r}
bike <- bike |> 
  mutate(date = dmy(date), # fix dates
         season = factor(season), # set as factors
         holiday = factor(holiday),
         functioning = factor(functioning))  
  
```

- Create some summary statistics

```{r}
bike |>
  group_by(functioning) |>
  summarize(n.bike_count = n(),
            across(.cols = c(bike_count),
                   list("mean" = ~ mean(.x, na.rm = TRUE),
                        "median" = ~ median(.x, na.rm = TRUE),
                        "sd" = ~ sd(.x, na.rm = TRUE),
                        "Q1" = ~ quantile(.x, 0.25, na.rm = TRUE),
                        "Q3" = ~ quantile(.x, 0.75, na.rm = TRUE),
                        "min" = ~ min(.x, na.rm = TRUE),
                        "max" = ~ max(.x, na.rm = TRUE)),
                   .names = "{.fn}.{.col}")) |> # creates names from the variable and the statistic
  pivot_longer(
    cols = -functioning, # exclude functioning from the pivot
    names_to = c("Statistic", "Variable"),
    names_sep = "\\.") |>
  pivot_wider(
    names_from = Statistic,
    values_from = value) |>
  gt(groupname_col = "functioning", rowname_col = "Variable") |> # organizes the Variable column by functioning and removes the "Variable" title
  fmt_number(decimals = 1) |> # rounds everything to 1 decimals
  tab_header(
    title = "Bike Count: Mean, Median, SD, Q1, and Q3",
    subtitle = "Stratified by functioning")
```

- Note:  count is zero for all non-functioning observations.  Going forwards, lets subset to omit non-functioning observations

```{r}
bike_valid <- bike |>
  filter(functioning == "Yes")
```

- Summarize across hours.  Use the sum of bike_count, rain, and snow.  Use the mean of temp, hum, wind, vis, dewpt, solrad

```{r}
bike_day <- bike_valid |>
  group_by(date, season, holiday) |>
  summarize(
    across(.cols = c(bike_count, rain, snow),  ~ sum(.x, na.rm = TRUE)), # sum variables
    across(.cols = c(temp, hum, wind, vis, dewpt, solrad), ~ mean(.x, na.rm = TRUE)), # mean variables
    .groups = "drop" # used to silence the output message
  )
```

- Take a look at some summary stats and plots after summarizing across hours

```{r}
summary(bike_day)
```

```{r}
bike_day |>
  group_by(season) |>
  summarize(n.bike_count = n(),
            across(.cols = c(bike_count),
                   list("mean" = ~ mean(.x, na.rm = TRUE),
                        "median" = ~ median(.x, na.rm = TRUE),
                        "sd" = ~ sd(.x, na.rm = TRUE),
                        "Q1" = ~ quantile(.x, 0.25, na.rm = TRUE),
                        "Q3" = ~ quantile(.x, 0.75, na.rm = TRUE),
                        "min" = ~ min(.x, na.rm = TRUE),
                        "max" = ~ max(.x, na.rm = TRUE)),
                   .names = "{.fn}.{.col}")) |> # creates names from the variable and the statistic
  pivot_longer(
    cols = -season, # exclude season from the pivot
    names_to = c("Statistic", "Variable"),
    names_sep = "\\.") |>
  pivot_wider(
    names_from = Statistic,
    values_from = value) |>
  gt(groupname_col = "season", rowname_col = "Variable") |> # organizes the Variable column by season and removes the "Variable" title
  fmt_number(decimals = 1) |> # rounds everything to 1 decimals
  tab_header(
    title = "Bike Count: Mean, Median, SD, Q1, and Q3",
    subtitle = "Stratified by season")

# bike count is lowest in winter and highest in autumn and summer
```

```{r}
bike_day |>
  group_by(holiday) |>
  summarize(n.bike_count = n(),
            across(.cols = c(bike_count),
                   list("mean" = ~ mean(.x, na.rm = TRUE),
                        "median" = ~ median(.x, na.rm = TRUE),
                        "sd" = ~ sd(.x, na.rm = TRUE),
                        "Q1" = ~ quantile(.x, 0.25, na.rm = TRUE),
                        "Q3" = ~ quantile(.x, 0.75, na.rm = TRUE),
                        "min" = ~ min(.x, na.rm = TRUE),
                        "max" = ~ max(.x, na.rm = TRUE)),
                   .names = "{.fn}.{.col}")) |> # creates names from the variable and the statistic
  pivot_longer(
    cols = -holiday, # exclude holiday from the pivot
    names_to = c("Statistic", "Variable"),
    names_sep = "\\.") |>
  pivot_wider(
    names_from = Statistic,
    values_from = value) |>
  gt(groupname_col = "holiday", rowname_col = "Variable") |> # organizes the Variable column by holiday and removes the "Variable" title
  fmt_number(decimals = 1) |> # rounds everything to 1 decimals
  tab_header(
    title = "Bike Count: Mean, Median, SD, Q1, and Q3",
    subtitle = "Stratified by holiday")

#  bike count is higher on average on non-holidays
```

```{r}
# Check correlation across all numeric variables

bike_day |>
  select(where(is.numeric)) |>
  cor() |>
  round(2)

# temp, dewpt, and solrad have the strongest correlation with bike count
```


```{r}
# Bike Count vs Temp

g <- ggplot(bike_day,
            aes(x = temp, y = bike_count))
g + geom_point(alpha = 1, size = 0.9) + 
  annotate("text",
           x = -10, y = 30000,
           label = paste("r =", round(cor(bike_day$temp, bike_day$bike_count),2))) + # calculate correlation and place on graph
  labs(x = "Temp (C)",
       y = "Bike Count",
       title = "Bike Count vs Temp") +
  theme(plot.title = element_text(hjust = 0.5)) # Centers the title
```

```{r}
# Bike Count vs Rain

g <- ggplot(bike_day,
            aes(x = rain, y = bike_count))
g + geom_point(alpha = 1, size = 0.9) + 
  annotate("text",
           x = 50, y = 30000,
           label = paste("r =", round(cor(bike_day$rain, bike_day$bike_count),2))) + # calculate correlation and place on graph
  labs(x = "Rain (mm)",
       y = "Bike Count",
       title = "Bike Count vs Rain") +
  theme(plot.title = element_text(hjust = 0.5)) # Centers the title
```

```{r}
# Bike Count vs snow

g <- ggplot(bike_day,
            aes(x = snow, y = bike_count))
g + geom_point(alpha = 1, size = 0.9) + 
  annotate("text",
           x = 30, y = 30000,
           label = paste("r =", round(cor(bike_day$snow, bike_day$bike_count),2))) + # calculate correlation and place on graph
  labs(x = "Snow (cm)",
       y = "Bike Count",
       title = "Bike Count vs Snow") +
  theme(plot.title = element_text(hjust = 0.5)) # Centers the title
```

```{r}
# Rain vs Humidity

g <- ggplot(bike_day,
            aes(x = hum, y = rain))
g + geom_point(alpha = 1, size = 0.9) + 
  annotate("text",
           x = 40, y = 60,
           label = paste("r =", round(cor(bike_day$hum, bike_day$rain),2))) + # calculate correlation and place on graph
  labs(x = "Humidity (%)",
       y = "Rain (mm)",
       title = "Rain vs Humidity") +
  theme(plot.title = element_text(hjust = 0.5)) # Centers the title
```

```{r}
# Bike Count vs Season

g <- ggplot(bike_day)
g + geom_boxplot(aes(x = season, y = bike_count, fill = season)) +
  labs(x = "Season",
       y = "Bike Count",
       title = "Bike Count by Season") + 
  theme(plot.title = element_text(hjust = 0.5), # Centers the title
        legend.position = "none") # remove redundant legend
```

```{r}
# Bike Count vs Holiday

g <- ggplot(bike_day)
g + geom_boxplot(aes(x = holiday, y = bike_count, fill = holiday)) +
  labs(x = "Holiday",
       y = "Bike Count",
       title = "Bike Count by Holiday") + 
  theme(plot.title = element_text(hjust = 0.5), # Centers the title
        legend.position = "none") # remove redundant legend
```

## Split the data and create 10-fold CV

- Split the data 75/25 for training and test

```{r}
set.seed(10)

# split 75/25, stratifying by season
bike_split <- initial_split(bike_day, prop = 0.75, strata = season)
bike_train <- training(bike_split)
bike_test <- testing(bike_split)

# check counts and stratification of each
bike_train |> 
  group_by(season) |>
  summarize(count = n())
bike_test |> 
  group_by(season) |>
  summarize(count = n())
```

- Create 10 fold CV split

```{r}
bike_10_fold <- vfold_cv(bike_train, 10)
```


## Fit three different MLR models on the training data using 10 fold CV to find best model
